---
title: "Homework 3"
author: Eric Sun
output: github_document
---

Initial setup

```{r}
library(tidyverse)
library(p8105.datasets)
library(janitor)

			 Knitr::opts_chunk$set(
				 Fig.width=6,
				 Fig.asp=.6,
				 Out.width="90%"
				 )
			 Theme_set(theme_minimal() + theme(legend.position="bottom"))
				 
			 Options(
				 Ggplot2.continuous.colour="viridis",
				 Ggplot2.continuous.fill="viridis")
			 Scale_colour_discrete=scale_color_viridis_d
			 Scale_fill_discrete=scale_fill_viridis_d

```

# Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>%
  count(aile %>%
          arrange(desc(n)))
```

Let's make a plot
```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Let's make a table
```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

Apples vs ice cream
```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

#Problem 2

Load, tidy, wrangle data; include a weekday vs weekend variable; and encode data with reasonable variable classes
```{r}
accel_df = read_csv("./data/accel_data.csv") %>%
  pivot_longer(activity.1:activity.1440, names_to = "minute", names_prefix="activity.", values_to = "count") %>%
  mutate(day=factor(day,c("Monday", "Tuesday", "Wednesday", "Thursday","Friday","Saturday","Sunday"))) %>%
  mutate(weekend=case_when(
    day %in% c("Saturday","Sunday") ~ "TRUE",
    day %in% c("Monday", "Tuesday", "Wednesday", "Thursday","Friday") ~ "FALSE")) %>%
  mutate(weekend=as.logical(weekend))
accel_df
```

The accel_df database contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. The variables describe the timing of the observations and the amount of activity for each observation.

